Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-27 prop_letter.tex
prop_letter.tex \section{Gleason} Updating...

Our framework allows us to generalize the above theorem to Born QIVPMs~$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$:
The collection of states consistent $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
gets closer and closer to $\rho$ as the intervals get narrower and
narrower, eventually collapsing to $\rho$ when the intervals become
infinitely precise real numbers covering the real line. To state this
result, we need to introduce a notion of distance between two states~$\rho$
and $\rho'$ as $\left\Vert \rho-\rho'\right\Vert =\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$~\citep{544199}.
The following theorem that is obtained by replacing the QIVPM~$\bar{\mu}$
in Thm.~\ref{cor:Gleason's} by a Born QIVPM~$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
and estimate the states consistent with it.

@@

\noindent If a state~$\rho'$ is consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
the difference between the Born quantum probabilities on any 1-dimensioal
projector~$\proj{\psi}$, $\left|\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)-\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)\right|$,
must be bounded by $1/n$ which implies $\left\Vert \rho-\rho'\right\Vert \le\frac{1}{n}$
by definition. As a concrete example of Thm.~\ref{thm:Finite-precision-Gleason},
since $\rho_{1}\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}},\events\right)$
and $\rho_{0}\in\coreBorn\left(\widehat{\iota}_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}},\events\right)$
which is easy to see on the commuting subspace of event~$\events_{C}$
according to Table~\ref{tab:Compute-the-QIVPM}, we have $\left\Vert \rho_{0}-\rho_{1}\right\Vert \le\frac{1}{2}$,
and this is indeed the case because $\left\Vert \rho_{0}-\rho_{1}\right\Vert =\frac{1}{24}$~\citep{HOSTunpublished}.

@@

\noindent Given an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $,
let $\mathscr{I}_{0}=\left\{ \necess,\imposs,\unknown\right\} $,
$\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$, $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$.
The function~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$ defined
in Table~\ref{tab:non-Born-QIVPM} can be verified as a QIVPM straightforwardly.
\begin{table}
\caption{\label{tab:non-Born-QIVPM}Define a non-Born QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$
on the three-dimensional Hilbert space.}

\begin{tabular}{cc}
\toprule 
\addlinespace
$P$ & $\bar{\mu}\left(P\right)$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
$\mathbb{0}$, $\proj{0}$, $\proj{\ps}$, $\proj{\ps'}$ & $\imposs$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\mathbb{1}$, $\mathbb{1}-\proj{0}$, $\mathbb{1}-\proj{\ps}$, $\mathbb{1}-\proj{\ps'}$ & $\necess$\tabularnewline\addlinespace
\midrule 
\addlinespace
All other projectors & $\unknown$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
Next we will prove $\coreBorn\left(\bar{\mu},\events\right)$ is an
empty set by contradiction. Suppose there is a state $\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}\in\coreBorn\left(\bar{\mu},\events\right)$,
where $\sum_{j=1}^{N}q_{j}=1$ and $q_{j}>0$. Since $\mu_{\rho}^{\mathrm{B}}(P)\in\bar{\mu}(P)$
for any~$P$, the state~$\rho$ must satisfy: 
\begin{equation}
\mu_{\rho}^{\mathrm{B}}(\proj{0})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps'})=0\,,
\end{equation}
which implies $\ip{\phi_{j}}{0}=\ip{\phi_{j}}{\ps}=\ip{\phi_{j}}{\ps'}=0$
for all $j$. The last equation never holds proving the theorem. Thus,
there is no state consistent with the QIVPM on all projectors~$P\in\events$,
so Gleason's theorem cannot be extended to general QIVPMs.

Notice that if we restrict $\bar{\mu}$ on the commuting subspace
of events~$\events_{C}$, the state~$\frac{1}{2}\cdot\proj{1}+\frac{1}{2}\cdot\proj{2}\in\coreBorn\left(\bar{\mu},\events_{C}\right)$
so that $\coreBorn\left(\bar{\mu},\events_{C}\right)$ is non-empty.
In general, for any commuting subspace of events~$\events_{C}'$,
$\coreBorn\left(\bar{\mu},\events_{C}'\right)$ is non-empty according
to the classical Thm.~\ref{thm:Shapley}. The difference between
classical and quantum highlights the following fact: Although classically
we could build a theory and deal the finite-precision measurement
in error analysis, it is necessary to embed the finite-precision measurement
into the foundation of quantum theory.

>>>>

Our framework allows us to generalize the above theorem to Born QIVPMs~$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$:
The collection of states consistent $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
gets closer and closer to $\rho$ as the intervals get narrower and
narrower, eventually collapsing to $\rho$ when the intervals become
infinitely precise real numbers covering the real line. To state this
result, we need to introduce a notion of distance between two states~$\rho$
and $\rho'$ as $\left\Vert \rho-\rho'\right\Vert =\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$~\citep{544199}.
By replacing the QIVPM~$\bar{\mu}$ in Thm.~\ref{cor:Gleason's}
by a Born QIVPM~$\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$, we obtain
the following theorem that estimates the states consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$.

@@

\noindent If a state~$\rho'$ is consistent with $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$,
the difference between their Born quantum probabilities on any 1-dimensioal
projector~$\proj{\psi}$, $\left|\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)-\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)\right|$,
must be bounded by $1/n$ which implies $\left\Vert \rho-\rho'\right\Vert \le\frac{1}{n}$
by definition. For example, Table~\ref{tab:Compute-the-QIVPM} shows
that $\rho_{0}$ is consistent with $\iota_{2}\circ\mu_{\rho_{1}}^{\mathrm{B}}$
and $\rho_{1}$ is consistent with $\iota_{2}\circ\mu_{\rho_{0}}^{\mathrm{B}}$
both on every projectors in $\events_{C}$. Indeed, they are consistent
with each other on every projectors in $\events$ that implies $\left\Vert \rho_{0}-\rho_{1}\right\Vert \le\frac{1}{2}$
by Thm.\ref{thm:Finite-precision-Gleason}. This is true because the
exact value of $\left\Vert \rho_{0}-\rho_{1}\right\Vert $ is $\frac{1}{24}$~\citep{544199}.

@@

\noindent Given an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $,
let $\mathscr{I}_{0}=\left\{ \necess,\imposs,\unknown\right\} $,
$\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$, $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$.
The function~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$ defined
in Table~\ref{tab:non-Born-QIVPM} can be verified as a QIVPM straightforwardly.
\begin{table}
\caption{\label{tab:non-Born-QIVPM}Define a non-Born QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{0}$
on the three-dimensional Hilbert space. The projectors mapping to
the same interval is listed in the column labeled by $P$ and separated
by commas.}

\begin{tabular}{cc}
\toprule 
\addlinespace
$P$ & $\bar{\mu}\left(P\right)$\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
$\mathbb{0}$, $\proj{0}$, $\proj{\ps}$, and $\proj{\ps'}$ & $\imposs$\tabularnewline\addlinespace
\midrule 
\addlinespace
$\mathbb{1}$, $\mathbb{1}-\proj{0}$, $\mathbb{1}-\proj{\ps}$, and
$\mathbb{1}-\proj{\ps'}$ & $\necess$\tabularnewline\addlinespace
\midrule 
\addlinespace
All other projectors & $\unknown$\tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
Next we will prove $\coreBorn\left(\bar{\mu},\events\right)$ is an
empty set by contradiction. Suppose there is a state $\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}\in\coreBorn\left(\bar{\mu},\events\right)$,
where $\sum_{j=1}^{N}q_{j}=1$ and $q_{j}>0$. Since $\mu_{\rho}^{\mathrm{B}}(P)\in\bar{\mu}(P)$
for any~$P$, the state~$\rho$ must satisfy: 
\begin{equation}
\mu_{\rho}^{\mathrm{B}}(\proj{0})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps'})=0\,,
\end{equation}
which implies $\ip{\phi_{j}}{0}=\ip{\phi_{j}}{\ps}=\ip{\phi_{j}}{\ps'}=0$
for all $j$. The last equation never holds proving the theorem. Thus,
there is no state consistent with the QIVPM on all projectors~$P\in\events$,
so Gleason's theorem cannot be extended to general QIVPMs.

Notice that if we restrict $\bar{\mu}$ on the commuting subspace
of events~$\events_{C}$, the state~$\frac{1}{2}\cdot\proj{1}+\frac{1}{2}\cdot\proj{2}$
is consistent with $\bar{\mu}$ so that $\coreBorn\left(\bar{\mu},\events_{C}\right)$
is non-empty. In general, for any commuting subspace of events~$\events_{C}'$,
$\coreBorn\left(\bar{\mu},\events_{C}'\right)$ is non-empty according
to the classical Thm.~\ref{thm:Shapley}. The difference between
classical and quantum highlights the following fact: Although classically
we could build a theory and deal the finite-precision measurement
in error analysis, it is necessary to embed the finite-precision measurement
into the foundation of quantum theory.
